Binning Strategies and Related Loss for Large Data

### 2015-04-15 Paper.Rnw
section 5.2 comments

The reduced binned data from spatial binning contains the center and count information for all bins.  The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot. The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities \citep{colorperception}.

>>>>

The reduced binned data from spatial binning contains the center and count information for all bins. 
\hh{This sounds like it is set in stone - it's not. We can define which information is kept for each of the bins. Make this an assumption rather than a statement and define what kind of center you are talking about - visual or numeric. }

The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot.
The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities \citep{colorperception}.

\hh{XXX I don't think that figure 12 is doing its job - and for that matter, we should talk about what its job is. There are too many different things going on in it, and the figure might feed better into your argument when things are separated out. }

\hh{XXX Let's try a first picture with a continuous scale and three discrete binned frequencies of four, seven, and ten bins. This should drive home the message that Healey is right, and we can't distinguish more than seven shades anyway, i.e. we don't see a difference between the continuously shaded plot and the one with discrete shades. This re-enforces our approach to treat color as a categorical variable, because we preceive it that way (otherwise why would we?)}

\hh{XXX figure 13 is a plot with a screaming message: never ever use a raw data mapping to show color. use quantiles or log frequencies. It doesn't scream that at the moment, so emphasize it in the text.}

\hh{XXX the rest of figure 12 should probably go into a figure 14 - to show that quantile binning and log frequency binning yield very similar results. We need to fix the legend. Right now they are atrocious (and I am aware that I am responsible for them, so I can say that :) )}
